### 2016-12-27 Chapters/experiments.tex
Review of Hidde's comment of Chap 3

Most studies of the growth of microorganisms have been done at steady state.
This is a reasonable and logical choice given that at steady state, the behavior of microorganisms is reproducible, which helps uncovering simple theories to apprehend their inherent complexity.
But while this condition can be easily achieved in the laboratory, microorganisms naturally spend very little time in steay state.
This motivated the construction of a new theoretical framework in Chapter~\ref{chap:theory} to study growth laws during growth transitions.

@@

Unfortunately, far more information is available on ribosome abundance at steady state than during growth transitions [refs].
The main reason for this bias on steady-state conditions is that, from an experimental point of view, growth transitions are hard to control and may depend on the history of the culture~[citation ? Bremmer ?].
However, as we discussed in Section~\ref{sec:discussion}, the few studies reported that have been reported seem to be consistent with our predictions, in the sense that they indicate that during growth transitions, the synthesis rate of ribomoses oscillates~\cite{gausing_regulation_1980,zengel_transcription_1986} and the ppGpp concentration manifests a rapid succession of increases and decreases~\cite{friesen_synthesis_1975,murray_control_2003}.
But we cannot decisively validate or disprove the model predictions from measurements that were carried out at the population level, where it is inherently hard to identify switching patterns.
What is needed for the verification of our predictions are dynamical single-cell measurements of the ribosome concentration during well-controlled growth transitions.
While the observation of an on-off strategy would show that simple models of the type introduced in Chapter~\ref{chap:theory} are instrumental in gaining a better understainding of microbial physiology, its refutation would also raise interesting questions. 
If maximization of growth rate and biomass accumulation have been retained by natural selection in \textit{E. coli}, which factors would prevent the cell from behaving optimally from a theoretical point of view?

The aim of this chapter is to measure \textit{in vivo}, during a growth transition, the resource allocation profile $\alpha (\cdot)$ and compare its dynamics with the gold standard established in Chapter~\ref{chap:theory}.
To this end, we have measured ribosomal abundance of \textit{E. coli} at the single-cell level during a nutrient upshift.
More precisely, we constructed a strain in which a fluorescent marker has been attached to a ribosomal subunit, thus allowing the \textit{in-vivo} monitoring of (changes in) the abundance of ribosomes.
In collaboration with Irina Mihalcescu of the Laboratoire Interdisciplinaire de Physique, we cultivated this \textit{E. coli} strain in a microfluidic device, allowing the long-term imaging of individual cells in well-controlled conditions, notably involving a classic upshift experiment from growth in minimal medium with acetate to minimal medium with glucose.
We then developed a Kalman smoothing method, in collaboration with Eugenio Cinquemain of Inria Grenoble - Rh\^{o}ne-Alpes, to estimate the variations of the growth rate and the relative ribosomal synthesis rate during this upshift, which allows $\alpha (\cdot)$ to be reconstructed.
While the preliminary results presented here do not allow a decisive validation of the expected behavior, among other things due to the experimental difficulties that were encountered and image analysis challenges, we believe they are promising as a first step towards the better understanding of global resource allocation during growth transitions.

@@

How can one measure the resource allocation variable in \textit{E. coli} cells?
There is no direct way to quantify $\alpha (\cdot)$ in an experiment, mostly because of its abstract nature.
To correctly reconstruct it, one has to know the value of every single term in the equations in which it appears.
In the model of Eqs~\ref{eq:pdef-exp}-\ref{eq:rdef-exp}, $\alpha$ only appears in Eq.~\ref{eq:rdef-exp}.
We can therefore identify $\alpha (t_k)$ for each time $t_k$ when the terms $\beta r(t_k)$, $\dot{r}(t_k)$, and $\frac{k_R \cdot p(t_k)}{K_R + p(t_k)} \cdot r(t_k)$ are known, or equivalently, the six individual components $r(t_k)$, $\dot{r}(t_k)$, $p(t_k)$, $k_R$, $K_R$, $\beta$ they are composed of.
Even when assuming that the parameters $k_R$, $K_R$, and $\beta$ are known, or at least easy to estimate independently, reconstructing the dynamics of $\alpha$ would require estimation of the (change in) concentration of the gene expression machinery ($r$, $\dot{r}$) and the precursor concentration $p$.

Are such measurements feasible in practice?
The abundance of ribosomes can be quantified, for instance, by measuring the total RNA of the cell~\cite{scott_interdependence_2010}, radioactive markers \cite{gausing_regulation_1980,zengel_transcription_1986}, fluorescent labels~\cite{bakshi_superresolution_2012}, or mass spectrometry techniques \cite{HwaMSBProteomics}.
However, only the use of fluorescent proteins complies with our need for real-time single-cell quantification.
Obtaining an estimate of precursor abundance is even more challenging, since there is no clear proxy for the totality of precursors in the cell and, while absolute quantification of all internal metabolites has been achieved in \textit{E. coli}~\cite{bennett_absolute_2009}, this method is also not suitable for the estimation of the precursor concentration dynamics in individual cells.

The problem of quantifying $\alpha$ can be simplified by applying a transformation of the model of Eqs~\ref{eq:pdef-exp}-\ref{eq:rdef-exp}.
Taking into account that by construction the growth rate $\mu(t)$ is given by
\[
	\mu (t) = \beta \frac{k_R \cdot p(t)}{K_R + p(t)} \cdot r(t),
\]
we can rewrite Eq.~\ref{eq:rdef-exp} as follows
\begin{equation}
\label{eq:dot_r}
\dot{r}(t) = \mu (t) \, \left(\frac{\alpha(t)}{\beta} - r(t) \right).
\end{equation}
The problem of estimating $\alpha(\cdot)$ is thus equivalent to the problem of estimating $r(\cdot)$, $\dot{r}(\cdot)$, $\mu (\cdot)$ and $\beta$.
If we are satisfied with estimating $\alpha$ up to a constant factor $\beta$, dynamically measuring the ribosome concentration and the growth rate would make $\alpha (\cdot) / \beta$ identifiable.

Given that fluorescent reporters of ribosomal proteins can provide information on both $r(\cdot)$ and $\dot{r}(\cdot)$ \cite{ValentinBioinformatics}, the above reformulation of the problem is promising for the purposes of our study.
Moreover, since the predicted bang-bang profiles in gene activity are easily hidden at the population level if cells are not synchronous, which is generally the case, expression of the ribosomal genes need to be monitored at high sampling density on the single-cell level, which is also possible with fluorescent reporters.
We therefore constructed an \textit{E. coli} strain in which the S2 ribosomal subunit, encoded by the gene \textit{rpsB}, has been tagged with a green fluorescent protein (GFP).
The strain with the chromosomal \textit{rpsB-gfp} fusion, inspired by previous work \cite{refBakshi}, produces fluorescent ribosomes that are quantifiable \textit{in vivo} in single cells while not affecting the growth rate (see Material and Methods~\ref{sec:methods_strain}).
The monitoring of single cells of this strain, by time-lapse fluorescence miscroscopy, thus makes it possible to estimate $r(\cdot)$ and $\dot{r}(\cdot)$ and reconstruct $\mu(t)$.

The growth conditions also need to be carefully chosen for this study.
In natural environments, for many bacteria one of the most frequently encountered growth transitions, or at least the transition on which selection is expected to operate, is the transition from stationary phase in a nutrient-deprived medium [see reference in recent PNAS paper by Balaban] to (exponential) growth after the renewed availability of nutrients.
The transition from stationary to exponential phase is difficult to study though as many hard-to-control parameters play a role, including the duration of the nutrient stress, the composition of the medium before growth arrest, and the accumulation of waste products in the medium \cite{reviewBremerDennisBiochimie}.
As a consequence, growth resumption only occurs after a lag phase of variable duration~[citation lag phase review to find], which is not included in the model.
For this reason, we decided to focus on a transition from exponential growth on a poor carbon source (acetate) to exponential growth on a rich carbon source (glucose) in a well-defined minimal medium.
By construction, this steady-state-to-steady-state transition requires a long acquisition time before and after the shift, respectively, to ensure that the cells have no memory of their physiological state before the transition and have enough time to reach the new steady state after the transition.
The use of the mother machine, a microfluidic device that was designed to sustain exponential growth for long periods of time, is therefore a good choice for this purpose.
It allows to maintain a constant flow of fresh medium and to effect fast transitions by switching the medium (see Material and Methods~\ref{sec:growth_condition} and \ref{sec:microflu}).
The complete experimental plan is summarized in Fig.~\ref{fig:experiment_schema}.

@@

An experiment following the plan of Fig.~\ref{fig:experiment_schema} was carried out, over a period of 40 h, but encountered a number of difficulties.
First, the motor displacing the microscope along the Z-axis intermittently stalled, thus deactivating the autofocus and the Z-compensation for the tilt of the [microfluidics???] device in the XY plane.
As a consequence, until the experimenter manually intervened, many of the images acquired were out of focus.
This is the reason why data points between -720 and -150 min are not exploitable, but fortunately the experiment was planned in such a way that when it really mattered, notably around the growth transition, someone was present to monitor the microscope.
Second, at around 7~h after the nutrient upshift, the bacteria started to die for reasons that are unknown, the most plausible hypothesis being a phage contamination.
Data analysis was therefore limited to 550 min after the upshift, when roughly 3/4 of the cells were still growing.
For these and other reasons, the experiment will have to be repeated (Section~\ref{sec:chap3_discussion}), but there was no time left for this in the framework of my PhD thesis.

The acquired data consisted of phase contrast and fluorescence images of 60 wells in total, in four different fields.
While quite a few image analysis programs have been reported in the literature~[citations schitcells + others], and some of these specifically address the mother machine design of the microfluidics device~[molysso + Wang et al], they all presented limitations when applied to our data.
For instance, the segmentation algorithms experienced difficulties due to the low resolution of the camera, and the phase contrast images had a superposed reflection band due to the microfluidics device. [LACK OF ALIGNMENT IS NOT VERY STRONG ARGUMENT, UNLESS IT CANNOT BE REPAIRED]
Moreover, the fluorescence density was concentrated in hot spots (Fig.~\ref{fig:data_acquisition}) and its intensity varied during the experiment. 
While this was expected, since ribosomes are localized in the cell poles~\cite{bakshi_superresolution_2012} and ribosomal abundance is known to vary with the growth rate~\cite{scott_interdependence_2010}, it complicated automatic segmentation on the fluorescence images.

@@

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/data_acquisition}
\caption{
\textbf{Results of data acquisition.}
We imaged 6 fields (X1Y1, X2Y1, X3Y1, X1Y2, X2Y2, X3Y2) each containing 15 wells (W0 to W14) for $\sim$40 hours.
From the 90 wells, 68 were suitable for further analysis (the others being empty, out of frame for some period of time, or plugged).
For all wells, data points are missing in the interval between -720 and -150 min because the camera was out of focus.
We stopped the analysis at 550 min, when about 3/4 of the bacteria were still growing, before the entire population died within a few hours for an unknown reason.
The image labelled "Raw image" is the last image analyzed for the highlighted well.
The bacterium on the left of this image was manually segmented by selecting two pixels at the poles on the fluorescence images (red cross).
A 6-pixel-wide rectangular mask was computed for each image, resulting in the "Segmented image" on the right.
Fluorescence intensities are expressed in Relative Fluorescence Units (RFU) on a 16-bit image (values from 0 to 65335) and were corrected for camera background, but not autofluorescence background (Material and Methods~\ref{sec:cell_segmentation} and Discussion~\ref{sec:chap3_discussion}).
The fluorescence intensity of the cell, expressed in units RFU/pixel/cell, was computed by dividing the sum of the fluorescence intensities of the pixels in the rectangle by the total number of pixels in the rectangle.
The cell length is the distance in pixels between the two poles (red line on the raw and segmented images).
The thick lines in green and black highlight the time-varying length and fluorescence density for the cell visible in the top images, labelled "X3Y2, W2".
The vertical dashed lines represent the time of the upshift from growth on acetate to growth on glucose.
}
\label{fig:data_acquisition}
\end{figure}

@@

As we presented in section~\ref{sec:exp_design}, our goal is to reconstruct the signal $\alpha (\cdot)$ during a growth transition.
By modeling \textit{E. coli} as a cylinder, the volume can be assumed directly proportional to the length of the cell.
If we assume a negligible background of autofluorescence (see Material and Methods~\ref{sec:cell_segmentation}), the fluorescence concentration in the cell can be assumed proportional to the ribosome concentration.
More precisely, based on the available data, we have the following measurement model at each time-point $t_k$, $0\leq k \leq N-1$:
\begin{eqnarray}
L(t_k) &= \lambda \cdot V(t_k) + \epsilon_k, \label{eq:mes_L}\\
F(t_k) &= \gamma \cdot r(t_k) + \eta_k, \label{eq:mes_F}
\end{eqnarray}
where $L(t_k)$ and $F(t_k)$ are the length and fluorescence density in units RFU/pixel measured at time $t_k$, respectively,  and $V(t_k)$  and $r(t_k)$) the corresponding actual volume and ribosome concentration at time $t_k$, respectively. ($\lambda$,$\gamma$) are unknown proportionality constants, and ($\epsilon_k$,$\eta_k$) measurement noise. [YOU MAY NEED TO SPECIFY THE ASSUMPTIONS ON THE NOISE DISTRIBUTION (0 MEAN, NORMAL?)]

As explained in Section~\ref{sec:exp_design}, reconstructing $\alpha (\cdot)$ requires information on $\mu (\cdot)$, $r(\cdot)$ and $\dot{r}(\cdot)$.
From Eq.~\ref{eq:mes_F}, we can obtain $r(\cdot)$ and $\dot{r}(\cdot)$ by smoothing interpolation and differentiation.
Similarly, $\mu (\cdot)$ can be derived from Eq.~\ref{eq:mes_L} as it is defined by
\[
\dot{V}(t) = \mu (t) \cdot V(t).
\]
However, smoothing interpolation is particularly sensible to the boundary conditions.
Since each division in the time series generates a new boundary condition, smoothing interpolation of our data is expected to be little robust.
For this reason, the growth rate of single cells is usually estimated by fitting an exponential function to the volume data between each cell~[citation Wang et al et Izard et al].
While this is suitable when bacteria are at steady state, it is not applicable during a growth transition, where we expect the growth rate to vary between successive divisions.
Other techniques are less sensible to the above problem~\cite{zulkower_robust_2015}, and we decided to use Kalman smoothing for our purpose~\cite{kailath_linear_2000,jazwinski_stochastic_2007}.

@@

The Kalman filtering problem can now be formulated as follows.
The state of a growing bacterial cell is described by the dynamical system
\begin{eqnarray}
\dot{r}(t) &=& \mu (t) \cdot \frac{\alpha(t)}{\beta} - \mu (t) \cdot r(t),\\
\dot{V}(t) &=& \mu (t) \cdot V(t),
\end{eqnarray}
with the following measurement model:
\begin{eqnarray}
L(t_k) &= \lambda \cdot V(t_k) + \epsilon_k,\\
F(t_k) &= \gamma \cdot r(t_k) + \eta_k,
\end{eqnarray}
where the variables are as defined as for Eqs~\ref{eq:dot_r}, \ref{eq:mes_L} and \ref{eq:mes_F}.
While $\beta$ can be obtained from the literature (Chapter~\ref{chap:theory}), $\lambda$ and $\gamma$ are unknown parameters that need to be estimated along with $\alpha (\cdot)$ and $\mu (\cdot)$.
Since we are satisfied with a qualitative reconstruction of $\alpha (\cdot)$, we can simplify the system by defining $r_\gamma = \gamma \cdot r$ and $V_\lambda = \lambda \cdot V$.
The dynamical system is consequently rewritten as
\begin{eqnarray}
\dot{r}_\gamma(t) &=& \mu (t) \cdot\frac{\gamma \cdot \alpha(t)}{\beta} - \mu (t) \cdot r_\gamma (t),\label{eq:r_prob}\\
\dot{V}_\lambda}t) &=& \mu (t) \cdot V_\lambda (t),\label{eq:V_prob}
\end{eqnarray}
with the new measurement model:
\begin{eqnarray}
F(t_k) &= r_\gamma (t_k) + \eta_k,\label{eq:F_prob}\\
L(t_k) &= V_\lambda(t_k) + \epsilon_k.\label{eq:L_prob}
\end{eqnarray}
The Kalman filtering problem thus becomes the reconstruction of $\gamma \alpha (\cdot) / \beta$ and $\mu (\cdot)$ from measurements $\left\{ F(t_0), F(t_1), ..., F(t_{N-1}) \right\}$ and $\left\{L(t_0), L(t_1), ..., L(t_{N-1}) \right\}$.

@@

In order to estimate the time-varying allocation of resources after the upshift, we use a similar Bayesian regularization approach as for the estimation of the growth rate.
We note $u(t) = \gamma \alpha (t) / \beta$ the resource allocation input that we wish to reconstruct.
The full model for the estimation of $u$ is given by
\begin{eqnarray}
\dot{r}_\gamma(t) &=& \hat{\mu} (t) \cdot u (t) - \hat{\mu} (t) \cdot r_\gamma (t), \nonumber\\
\dot{u}(t) &=& v(t),\label{eq:full_u_prob}\\
\dot{v}(t) &=& w(t),\nonumber
\end{eqnarray}
with the measurement model
\begin{equation*}
F(t_k) = r_\gamma (t_k) + \eta_k,\\
\end{equation*}
where $\hat{\mu}$ is the estimation of $\mu$ obtained in the previous section.
Contrary to the model used for the estimation of the growth rate, the system of Eq.~\ref{eq:full_u_prob} is linear.
Furthermore, we do not have to deal with discontinuities in the measurements, since the fluorescence density $F$ in the cells is continuous between mother and daughter cells (Fig.~\ref{fig:data_acquisition}).

>>>>

Most studies of the growth of microorganisms have been done at steady state.
This is a reasonable and logical choice given that at steady state, the behavior of microorganisms is reproducible, which helps uncovering simple theories to apprehend their inherent complexity.
But while this condition can be easily achieved in the laboratory, microorganisms naturally spend very little time in steady state.
This motivated the construction of a new theoretical framework in Chapter~\ref{chap:theory} to study growth laws during growth transitions.

@@

Unfortunately, far more information is available on ribosome abundance at steady state than during growth transitions~\cite{scott_interdependence_2010,gausing_regulation_1980}.
The main reason for this bias on steady-state conditions is that, from an experimental point of view, growth transitions are hard to control and may depend on the history of the culture~\cite{ng_damage_1962,dufrenne_effect_1997,shaw_effect_1967,mcmeekin_predictive_2002,cheroutre-vialette_application_2002}.
However, as we discussed in Section~\ref{sec:discussion}, the few studies that have been reported seem to be consistent with our predictions, in the sense that they indicate that during growth transitions, the synthesis rate of ribomoses oscillates~\cite{gausing_regulation_1980,zengel_transcription_1986} and the ppGpp concentration manifests a rapid succession of increases and decreases~\cite{friesen_synthesis_1975,murray_control_2003}.
But we cannot decisively validate or disprove the model predictions from measurements that were carried out at the population level, where it is inherently hard to identify switching patterns.
What is needed for the verification of our predictions are dynamical single-cell measurements of the ribosome concentration during well-controlled growth transitions.
While the observation of an on-off strategy would show that simple models of the type introduced in Chapter~\ref{chap:theory} are instrumental in gaining a better understanding of microbial physiology, its refutation would also raise interesting questions. 
If maximization of growth rate and biomass accumulation have been retained by natural selection in \textit{E. coli}, which factors would prevent the cell from behaving optimally from a theoretical point of view?

The aim of this chapter is to measure \textit{in vivo}, during a growth transition, the resource allocation profile $\alpha (\cdot)$ and compare its dynamics with the gold standard established in Chapter~\ref{chap:theory}.
To this end, we have measured ribosomal abundance of \textit{E. coli} at the single-cell level during a nutrient upshift.
More precisely, we constructed a strain in which a fluorescent marker has been attached to a ribosomal subunit, thus allowing the \textit{in-vivo} monitoring of (changes in) the abundance of ribosomes.
In collaboration with Irina Mihalcescu of the Laboratoire Interdisciplinaire de Physique, we cultivated this \textit{E. coli} strain in a microfluidic device, allowing the long-term imaging of individual cells in well-controlled conditions, notably involving a classic upshift experiment from growth in minimal medium with acetate to minimal medium with glucose.
We then developed a Kalman smoothing method, in collaboration with Eugenio Cinquemani of Inria Grenoble -- Rh\^{o}ne-Alpes, to reconstruct $\alpha (\cdot)$  from the estimates of the variations of the growth rate and the relative ribosomal synthesis rate.
While the preliminary results presented here do not allow a decisive validation of the expected behavior, among other things due to the difficulties that were encountered during the experiments and the image analysis, we believe they are promising as a first step towards the better understanding of global resource allocation during growth transitions.

@@

How can one measure the resource allocation variable in \textit{E. coli} cells?
There is no direct way to quantify $\alpha (\cdot)$ in an experiment, mostly because of its abstract nature.
To correctly reconstruct it, one has to know the value of every single term in the equations in which it appears.
In the model of Eqs~\ref{eq:pdef-exp}-\ref{eq:rdef-exp}, $\alpha$ only appears in Eq.~\ref{eq:rdef-exp}.
We can therefore identify $\alpha (t_k)$ for each time $t_k$ when the terms $\beta r(t_k)$, $\dot{r}(t_k)$, and $\frac{k_R \cdot p(t_k)}{K_R + p(t_k)} \cdot r(t_k)$ are known, or equivalently, the six individual components $r(t_k)$, $\dot{r}(t_k)$, $p(t_k)$, $k_R$, $K_R$, $\beta$ they are composed of.
Even when assuming that the parameters $k_R$, $K_R$, and $\beta$ are known, or at least easy to estimate independently, reconstructing the dynamics of $\alpha$ would require estimation of the (changes in) concentration of the gene expression machinery ($r$, $\dot{r}$) and the precursor concentration $p$.

Are such measurements feasible in practice?
The abundance of ribosomes can be quantified, for instance, by measuring the total RNA of the cell~\cite{scott_interdependence_2010}, or using radioactive markers \cite{gausing_regulation_1980,zengel_transcription_1986}, fluorescent labels~\cite{bakshi_superresolution_2012}, or mass spectrometry techniques~\cite{hui_quantitative_2015}.
However, only the use of fluorescent proteins complies with our need for real-time single-cell quantification.
Obtaining an estimate of precursor abundance is even more challenging, since there is no clear proxy for the totality of precursors in the cell and, while absolute quantification of all internal metabolites has been achieved in \textit{E. coli}~\cite{bennett_absolute_2009}, this method is also not suitable for a dynamical estimation in individual cells.

The problem of quantifying $\alpha (\cdot)$ can be simplified by applying a transformation of the model of Eqs~\ref{eq:pdef-exp}-\ref{eq:rdef-exp}.
Taking into account that by construction the growth rate $\mu(t)$ is given by
\[
	\mu (t) = \beta \frac{k_R \cdot p(t)}{K_R + p(t)} \cdot r(t),
\]
we can rewrite Eq.~\ref{eq:rdef-exp} as follows
\begin{equation}
\label{eq:dot_r}
\dot{r}(t) = \mu (t) \, \left(\frac{\alpha(t)}{\beta} - r(t) \right).
\end{equation}
The problem of estimating $\alpha(\cdot)$ is thus equivalent to the problem of estimating $r(\cdot)$, $\dot{r}(\cdot)$, $\mu (\cdot)$ and $\beta$.
If we are satisfied with estimating $\alpha$ up to a constant factor $\beta$, dynamically measuring the ribosome concentration and the growth rate would make $\alpha (\cdot) / \beta$ identifiable.

Given that fluorescent reporters of ribosomal proteins can provide information on both $r(\cdot)$ and $\dot{r}(\cdot)$ \cite{zulkower_robust_2015}, the above reformulation of the problem is promising for the purposes of our study.
Moreover, since the predicted bang-bang profiles in gene activity are easily hidden at the population level if cells are not synchronous, which is generally the case, expression of the ribosomal genes need to be monitored at high sampling density on the single-cell level, which is also possible with fluorescent reporters.
Inspired by previous work~\cite{bakshi_superresolution_2012}, we therefore constructed an \textit{E. coli} strain in which the S2 ribosomal subunit, encoded by the gene \textit{rpsB}, has been tagged with a green fluorescent protein (GFP).
The strain with the chromosomal \textit{rpsB-gfp} fusion produces fluorescent ribosomes that are quantifiable \textit{in vivo} in single cells while not affecting the growth rate (see Material and Methods~\ref{sec:methods_strain} and~\cite{bakshi_superresolution_2012}).
Monitoring single cells of this strain, by time-lapse fluorescence microscopy, thus makes it possible to estimate $r(\cdot)$ and $\dot{r}(\cdot)$ and reconstruct $\mu(\cdot)$.

The growth conditions also need to be carefully chosen for this study.
In natural environments, for many bacteria one of the most frequently encountered growth transitions, or at least the transition on which selection is expected to operate, is the transition from stationary phase in a nutrient-deprived medium~\cite{gefen_direct_2014} to (exponential) growth after the renewed availability of nutrients.
The transition from stationary to exponential phase is difficult to study though as many hard-to-control parameters play a role, including the duration of the nutrient stress, the composition of the medium before growth arrest, and the accumulation of waste products in the medium~\cite{ehrenberg_mediumdependent_2012}.
As a consequence, growth resumption only occurs after a lag phase of variable duration~\cite{ng_damage_1962,dufrenne_effect_1997,shaw_effect_1967,mcmeekin_predictive_2002,cheroutre-vialette_application_2002}, which is not included in the model.
For this reason, we decided to focus on a transition from exponential growth on a poor carbon source (acetate) to exponential growth on a rich carbon source (glucose) in a well-defined minimal medium.
By construction, this steady-state-to-steady-state transition requires a long acquisition time before and after the shift, respectively, to ensure that the cells have no memory of their physiological state before the transition and have enough time to reach the new steady state after the transition.
The use of the mother machine, a microfluidic device that was designed to sustain exponential growth for long periods of time, is therefore a good choice for this purpose.
It allows to maintain a constant flow of fresh medium and to effect fast transitions by switching the medium (see Material and Methods~\ref{sec:growth_condition} and \ref{sec:microflu}).
The complete experimental plan is summarized in Fig.~\ref{fig:experiment_schema}.

@@

An experiment following the plan of Fig.~\ref{fig:experiment_schema} was carried out, over a period of 40~h, but encountered a number of difficulties.
First, the motor displacing the microscope along the Z-axis intermittently stalled, thus deactivating the autofocus and the Z-compensation for the tilt of the microfluidic device in the XY plane.
As a consequence, until the experimenter manually intervened, many of the images acquired were out of focus.
This is the reason why data points between -720 and -150 min are not exploitable, but fortunately the experiment was planned in such a way that when it really mattered, notably around the growth transition, someone was present to monitor the microscope.
Second, at around 7~h after the nutrient upshift, the bacteria started to die for reasons that are unknown, the most plausible hypothesis being a phage contamination (see Discussion~\ref{sec:chap3_discussion}).
Data analysis was therefore limited to 550 min after the upshift, when roughly 3/4 of the cells were still growing.
For these and other reasons (see Section~\ref{sec:chap3_discussion}), the experiment will have to be repeated, but there was no time left for this in the framework of my PhD thesis.

The acquired data consisted of phase contrast and fluorescence images of 60 wells in total, in four different fields.
While quite a few image analysis programs have been reported in the literature~\cite{klein_tlm-tracker:_2012,wang_image_2010,young_measuring_2012,paintdakhi_oufti:_2016}, and some of these specifically address the mother machine design of the microfluidic device~\cite{wang_robust_2010,sachs_image-based_2016}, they all presented limitations when applied to our data.
For instance, the segmentation algorithms experienced difficulties due to the low resolution of the camera, and the phase contrast images had a superposed reflection band due to the microfluidic device.
Moreover, the fluorescence density was concentrated in hot spots (Fig.~\ref{fig:data_acquisition}) and its intensity varied during the experiment. 
While this was expected, since ribosomes are localized in the cell poles~\cite{bakshi_superresolution_2012} and ribosomal abundance is known to vary with the growth rate~\cite{scott_interdependence_2010}, it complicated automatic segmentation on the fluorescence images.

@@

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/data_acquisition}
\caption{
\textbf{Results of data acquisition.}
We imaged 6 fields (X1Y1, X2Y1, X3Y1, X1Y2, X2Y2, X3Y2) each containing 15 wells (W0 to W14) for $\sim$40 hours.
From the 90 wells, 68 were suitable for further analysis (the others being empty, out of frame for some period of time, or plugged).
For all wells, data points are missing in the interval between -720 and -150 min because the camera was out of focus.
We stopped the analysis at 550 min, when about 3/4 of the bacteria were still growing, before the entire population died within a few hours for an unknown reason.
The image labelled "Raw image" is the last image analyzed for the highlighted well.
The bacterium on the left of this image was manually segmented by selecting two pixels at the poles on the fluorescence images (red cross).
A 6-pixel-wide rectangular mask was computed for each image, resulting in the "Segmented image" on the right.
Fluorescence intensities are expressed in Relative Fluorescence Units (RFU) on a 16-bit image (values from 0 to 65335) and were corrected for camera background, but not autofluorescence background (Material and Methods~\ref{sec:cell_segmentation} and Discussion~\ref{sec:chap3_discussion}).
The fluorescence intensity of the cell, expressed in units RFU/pixel/cell, was computed by dividing the sum of the fluorescence intensities of the pixels in the rectangle by the total number of pixels in the rectangle.
The cell length is the distance in pixels between the two poles (red line).
The thick lines in green and black highlight the time-varying length and fluorescence density for the cell visible in the top images, labelled "X3Y2, W2".
The vertical dashed lines represent the time of the upshift from growth on acetate to growth on glucose.
}
\label{fig:data_acquisition}
\end{figure}

@@

As we presented in section~\ref{sec:exp_design}, our goal is to reconstruct the signal $\alpha (\cdot)$ during a growth transition.
By modeling \textit{E. coli} as a cylinder, the volume can be assumed directly proportional to the length of the cell.
If we assume a negligible background of autofluorescence (see Material and Methods~\ref{sec:cell_segmentation}), the fluorescence concentration in the cell can be assumed proportional to the ribosome concentration.
More precisely, based on the available data, we have the following measurement model at each time-point $t_k$, $0\leq k \leq N-1$:
\begin{eqnarray}
L(t_k) &= \lambda \cdot V(t_k) + \epsilon_k, \label{eq:mes_L}\\
F(t_k) &= \gamma \cdot r(t_k) + \eta_k, \label{eq:mes_F}
\end{eqnarray}
where $L(t_k)$ and $F(t_k)$ are the length and fluorescence density in units RFU/pixel measured at time $t_k$, respectively,  and $V(t_k)$  and $r(t_k)$ the corresponding actual volume and ribosome concentration at time $t_k$, respectively. ($\lambda$,$\gamma$) are unknown proportionality constants, and ($\epsilon_k$,$\eta_k$) uncorrelated sequence of measurement noise assumed normally distributed with mean 0.

As explained in Section~\ref{sec:exp_design}, reconstructing $\alpha (\cdot)$ requires information on $\mu (\cdot)$, $r(\cdot)$ and $\dot{r}(\cdot)$.
From Eq.~\ref{eq:mes_F}, we can obtain $r(\cdot)$ and $\dot{r}(\cdot)$ by smoothing interpolation and differentiation.
Similarly, $\mu (\cdot)$ can be derived from Eq.~\ref{eq:mes_L} as it is defined by
\[
\dot{V}(t) = \mu (t) \cdot V(t).
\]
However, smoothing interpolation is particularly sensible to the boundary conditions.
Since each division in the time series generates a new boundary condition, smoothing interpolation of our data is expected to be little robust.
For this reason, the growth rate of single cells is usually estimated by fitting an exponential function to the volume data between each cell~\cite{wang_robust_2010,izard_synthetic_2015}.
While this is suitable when bacteria are at steady state, it is not applicable during a growth transition, where we expect the growth rate to vary between successive divisions.
Other techniques are less sensible to the above problem~\cite{zulkower_robust_2015}, and we decided to use Kalman smoothing for our purpose~\cite{kailath_linear_2000,jazwinski_stochastic_2007}.

@@

The Kalman filtering problem can now be formulated as follows.
The state of a growing bacterial cell is described by the dynamical system
\begin{eqnarray}
\dot{r}(t) &=& \mu (t) \cdot \frac{\alpha(t)}{\beta} - \mu (t) \cdot r(t),\\
\dot{V}(t) &=& \mu (t) \cdot V(t),
\end{eqnarray}
with initial conditions $r(0) = r_0$, $V(0) = V_0$, and the following measurement model:
\begin{eqnarray}
L(t_k) &= \lambda \cdot V(t_k) + \epsilon_k,\\
F(t_k) &= \gamma \cdot r(t_k) + \eta_k,
\end{eqnarray}
where the variables are as defined as for Eqs~\ref{eq:dot_r}, \ref{eq:mes_L} and \ref{eq:mes_F}.
While $\beta$ can be obtained from the literature (Chapter~\ref{chap:theory}), $\lambda$ and $\gamma$ are unknown parameters that need to be estimated along with $\alpha (\cdot)$ and $\mu (\cdot)$.
Since we are satisfied with a qualitative reconstruction of $\alpha (\cdot)$, we can simplify the system by defining $r_\gamma = \gamma \cdot r$ and $V_\lambda = \lambda \cdot V$.
The dynamical system is consequently rewritten as
\begin{eqnarray}
\dot{r}_\gamma(t) &=& \mu (t) \cdot\frac{\gamma \cdot \alpha(t)}{\beta} - \mu (t) \cdot r_\gamma (t),\label{eq:r_prob}\\
\dot{V}_\lambda (t) &=& \mu (t) \cdot V_\lambda (t),\label{eq:V_prob}
\end{eqnarray}
with initial conditions $r_\gamma (0) = \gamma r_0$, $V_\lambda (0) = \lambda V_0$, and the new measurement model:
\begin{eqnarray}
F(t_k) &= r_\gamma (t_k) + \eta_k,\label{eq:F_prob}\\
L(t_k) &= V_\lambda(t_k) + \epsilon_k.\label{eq:L_prob}
\end{eqnarray}
The Kalman filtering problem thus becomes the reconstruction of $\gamma \alpha (\cdot) / \beta$ and $\mu (\cdot)$ from measurements $\left\{ F(t_0), ..., F(t_{N-1}) \right\}$ and $\left\{L(t_0), ..., L(t_{N-1}) \right\}$.

@@

In order to estimate the time-varying allocation of resources after the upshift, we use a similar Bayesian regularization approach as for the estimation of the growth rate.
We note $u(t) = \gamma \alpha (t) / \beta$ the resource allocation input that we wish to reconstruct.
With $\hat{\mu}$ the estimation of $\mu$ obtained in the previous section, the full model for the reconstruction of $u (\cdot)$ is given by
\begin{eqnarray}
\dot{r}_\gamma(t) &=& \hat{\mu} (t) \cdot u (t) - \hat{\mu} (t) \cdot r_\gamma (t), \nonumber\\
\dot{u}(t) &=& v(t),\label{eq:full_u_prob}\\
\dot{v}(t) &=& w(t),\nonumber
\end{eqnarray}
with the measurement model
\begin{equation}
F(t_k) = r_\gamma (t_k) + \eta_k,\\
\end{equation}
and the initial conditions
\[
r_\gamma (0) = r_{\gamma,0}, \;\;\;\; u (0) = u_0 , \;\;\;\; v(0) = v_0,
\]
that may themselves be Gaussian random variables with some given statistics.
Contrary to the model used for the estimation of the growth rate, the system of Eq.~\ref{eq:full_u_prob} is linear.
Furthermore, we do not have to deal with discontinuities in the measurements, since the fluorescence density $F$ in the cells is continuous between mother and daughter cells (Fig.~\ref{fig:data_acquisition}).
